---
title: "<b>Electric Vehicle Charging Devices (Scotland)</b>"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float: true
    code_folding: "hide"
  pdf_document: default
---

```{r, echo = FALSE, warning = FALSE, message = FALSE}
library(tidyverse)
library(sf)
library(janitor)
library(leaflet)
library(htmlwidgets)
library(gt)
library(gghighlight)
```

```{r, echo = FALSE, warning = FALSE, message = FALSE}

ncr_data <- read_csv("data/raw_data/national-charge-point-registry.csv") %>% 
  clean_names()
  
postcodes <- read_csv("data/raw_data/scotland_postcodes.csv") %>% 
  clean_names() %>% 
  select(postcode, district)

population <- read_csv("data/raw_data/council-area-profiles-dataset-with-copyright_population-estimates.csv") %>% 
  clean_names()

```

```{r, echo = FALSE, warning = FALSE, message = FALSE}

ncr_council_summary <- ncr_data %>%
  filter(charge_device_status == "In service") %>% 
  select(charge_device_id, postcode) %>% 
  inner_join(postcodes, "postcode") %>%
  mutate(district = recode(district, "Na h-Eileanan Siar" = "Eilean Siar")) %>% 
  count(district)
  
```

```{r, echo = FALSE, warning = FALSE, message = FALSE}
population_summary <- population %>%
  filter(year == "2021",
         council_area != "Scotland") %>%
  mutate(council_area = recode(council_area, "Na h-Eileanan Siar" = "Eilean Siar")) %>%
  group_by(council_area) %>% 
  summarise(total_population = sum(population))
```

```{r, echo = FALSE, warning = FALSE, message = FALSE}
ncr_council_summary_pop <- ncr_council_summary %>% 
  inner_join(population_summary, by=c("district" = "council_area")) %>% 
  mutate(devices_per_hundred_thousand = n / (total_population / 100000))
```

*Disclaimer: This report is intended as a proof ability to produce technical reports and visualisations using relevant data. Additional data sources and investigation is required in order to make a final estimate on the numbers of chargepoints in each Local Authority.*

# Introduction

Based on data from the [National Chargepoint Registry](https://www.gov.uk/guidance/find-and-use-data-on-public-electric-vehicle-chargepoints) the following report provides insight into the number of charging devices relative to the population size and the total number of devices in Scotland's Local Authorities.

Data was sources from the National Chargepoint Registry in *July 2023* and the post-codes provided for each device used to map these to the appropriate Local Authority. Only exact post-code matches were accepted.

***

# Devices Per 100,000 of Population

<br>

<div class = "row">
  <div class = "col-md-6">
  
<!--html_preserve-->
<iframe src = "maps/fig_2.html" width="425" height="500" frameBorder="0"> </iframe>
<!--/html_preserve-->

</div>
<div class = "col-md-6">

```{r, echo = FALSE, warning = FALSE, message = FALSE}
ncr_council_summary_pop %>% 
  select(district, devices_per_hundred_thousand) %>%
  mutate(devices_per_hundred_thousand = floor(devices_per_hundred_thousand)) %>%
  arrange(desc(devices_per_hundred_thousand)) %>%
  gt() %>%
  cols_label(
    district = md("**Local Authority**"),
    devices_per_hundred_thousand = md("**Devices Per 100,000**")) %>% 
  opt_interactive(use_sorting = FALSE, 
                  use_search = TRUE, 
                  use_compact_mode = TRUE) %>% 
  tab_options(column_labels.font.weight = "bold", table.align = "left") %>% 
  data_color(columns = devices_per_hundred_thousand, palette = "YlOrRd")
```

</div>
</div>

<br>

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">

- The Island regions of Scotland appear to have some of the highest number of devices per 100,000 population. However, relatively low population totals in these regions may contribute.

- Of the mainland regions, Dumfries and Galloway appear to have the highest number of charging devices per 100,000 population.

</div>

***

# Total Devices

<br>

<div class = "row">
  <div class = "col-md-6">

<!--html_preserve-->
<iframe src = "maps/fig_1.html" width="425" height="500" frameBorder="0"> </iframe>
<!--/html_preserve-->

</div>

<div class = "col-md-6">

```{r, echo = FALSE, warning = FALSE, message = FALSE}
ncr_council_summary_pop %>% 
  select(district, n) %>% 
  arrange(desc(n)) %>%
  gt() %>%
  cols_label(
    district = md("**Local Authority**"),
    n = md("**Charging Devices**")) %>% 
  opt_interactive(use_sorting = FALSE, 
                  use_search = TRUE, 
                  use_compact_mode = TRUE) %>% 
  tab_options(column_labels.font.weight = "bold", table.align = "left") %>% 
  data_color(columns = n, palette = "YlOrRd")
```


</div>
</div>

<br>

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">

- **Glasgow** has the highest overall number of charging devices; however, this is likely expected as it also has the highest total population.

- Notably, the region with the **2nd** highest number of devices is **Dumfries and Galloway** despite being ranked **13th** in population size.

</div>

***

# Trends Over Time

<br>

```{r, echo = FALSE, warning = FALSE, message = FALSE, fig.align = 'center'}

ncr_data %>%
  filter(charge_device_status == "In service") %>%
  inner_join(postcodes, "postcode") %>%
  mutate(district = recode(district, "Na h-Eileanan Siar" = "Eilean Siar"),
         date_created = as.POSIXct(date_created),
         date_created = year(date_created)) %>%
  filter(date_created > 2017,
         date_created < 2023) %>% 
  group_by(date_created, district) %>% 
  summarise(annual_new_by_district = n()) %>%
  ggplot(aes(x = date_created, y = annual_new_by_district, colour = district)) +
  geom_line(size = 1.5) +
  geom_point(size = 3) +
  gghighlight(max(annual_new_by_district) > 60) +
  labs(
    x = "\n Year",
    y = "New Devices Added \n",
    title = "Devices Added To National Chargepoint Registry: 2018-2022",
    subtitle = "By Local Authority",
    colour = "Local Authority") +
  theme(panel.background = element_rect(fill = "white"),
        panel.grid = element_line(colour = "grey90"),
        plot.title = element_text(size = 15, face = "bold"),
        plot.subtitle = element_text(size = 10))
```

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">

- Following the findings above, we can see that **Dumfries and Galloway** added **68** new devices in 2022. This is the largest number of new devices added by any region in one year over the last five years. The previous highest number was **29** added in Glasgow in 2021.

</div>

***

# Conclusions

<style>
div.yellow {background-color:#fcf8c0; border-radius: 5px; padding: 20px;}

</style>
<div class = "yellow">

- **Dumfries and Galloway** is notable for having a significant amount of charging devices per 100,000 population, as well as a high total number of charging devices.

- **Dumfries and Galloway** added more new devices in 2022 than any region has done in a previous year (68).

- Further work should be completed to understand the nature of the increase in new devices added in Dumfries and Galloway in 2022. For example, were these all newly installed devices, OR was there a delay in adding devices to the register from previous years? If these were all new devices, consideration should be given to the strategy and initiatives utilised.

</div>

***

# Data Sources

- Source: National Chargepoint Registry / Open Government Licence v3.0 / Crown Copyright 2023
https://www.gov.uk/guidance/find-and-use-data-on-public-electric-vehicle-chargepoints

- Source: National Records of Scotland: Council Area Profiles / Open Government Licence v3.0 / Crown Copyright 2023
https://www.nrscotland.gov.uk/statistics-and-data/statistics/stats-at-a-glance/council-area-profiles

- spatialdata.gov.scot / Open Government Licence v3.0 / Crown Copyright 2023
https://spatialdata.gov.scot/geonetwork/srv/api/records/1cd57ea6-8d6e-412b-a9dd-d1c89a80ad62

- doogal.co.uk / Public Domain 
https://www.doogal.co.uk/PostcodeDownloads